---
title: "Investigating Hungarian Agricultural Subsidies Data"
subtitle: "CEU Business Analytics MSc"
author: "Gergo Szekely"
date: "2020 May"
fig_height: 7
urlcolor: blue
output:
  pdf_document:
    fig_height: 3.1
    fig_width: 5.1
    keep_tex: yes
    number_sections: yes
    toc: true
  html_document:
    df_print: paged
    toc: true
    theme: united
  word_document: default
header-includes: \usepackage{dcolumn} \usepackage{float} \floatplacement{figure}{H}
---


```{r setup, include=FALSE, echo = FALSE, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, fig.pos = 'H')
rm(list = ls())
sink("/dev/null")

list.of.packages <-c("tidyverse",
                     "ggplot2",
                     "data.table",
                     "plot3D",
                     "arsenal",
                     "forcats",
                     "viridis"
                     )


new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[,"Package"])]
if(length(new.packages)>0) {install.packages(new.packages)}
lapply(list.of.packages, require, character.only = TRUE)

dir <- getwd()
data_in <- paste0(dir,"/../output/")
data_out <- paste0(dir,"/../imgs/")

agrar_funds <- readr::read_csv(paste0(data_in, "agrar_funds.csv"))
agrar_funds$fund_id <- as.integer(agrar_funds$fund_id)

agrar_wins <- readr::read_csv(paste0(data_in, "agrar_wins.csv"))
agrar_wins$winner_id <- as.integer(agrar_wins$winner_id)
agrar_wins$fund_id <- as.integer(agrar_wins$fund_id)

agrar_winners <- readr::read_csv(paste0(data_in, "agrar_winners.csv"),
                                   col_types = cols(
                                     name = col_character(),
                                      gender = col_character(),
                                      address = col_character(),
                                      formatted_address = col_character(),
                                      is_firm = col_double(),
                                      latitude = col_double(),
                                      longitude = col_double(),
                                      settlement_id = col_double(),
                                      winner_id = col_double()
                                     )
                                 )
agrar_winners$settlement_id <- as.integer(agrar_winners$settlement_id)
agrar_winners$winner_id <- as.integer(agrar_winners$winner_id)

agrar_settlements <- readr::read_csv(paste0(data_in, "agrar_settlements.csv"))
agrar_settlements$settlement_id <- as.integer(agrar_settlements$settlement_id)

agrar_funds$land_based <- as.factor(agrar_funds$land_based)
agrar_wins$year <- as.factor(agrar_wins$year)
agrar_funds$reason <- as.factor(agrar_funds$reason)
agrar_funds$program <- as.factor(agrar_funds$program)
agrar_funds$source <- as.factor(agrar_funds$source)
agrar_settlements$settlement_type <- factor(agrar_settlements$settlement_type,                                     ordered = TRUE, levels = c("Budapest","megyeszékhely","járás székhely","város","nagyközség","község"))
agrar_winners$is_firm <- as.factor(agrar_winners$is_firm)
agrar_winners$gender <- as.factor(agrar_winners$gender)
agrar_settlements$settlement_type <- as.factor(agrar_settlements$settlement_type)

agrar_full <- agrar_wins %>%
  inner_join(agrar_winners, by = "winner_id") %>%
  inner_join(agrar_settlements, by = "settlement_id") %>%
  inner_join(agrar_funds, by = "fund_id")
  
rm(agrar_wins)
rm(agrar_winners)
rm(agrar_funds)
rm(agrar_settlements)

agrar_full$winner_id <- NULL
agrar_full$fund_id <- NULL
agrar_full$settlement_id <- NULL

options(scipen=999)
set.seed(42)

sum_by_dummies <- agrar_full %>%
  select(amount,year,is_firm,land_based,settlement_type,source) %>%
  mutate(year = as.factor(year)) %>%
  group_by(year,is_firm,land_based,settlement_type,source) %>%
  summarise(amount=sum(amount))

sum_by_address <- agrar_full %>%
  select(year,zip,address,amount) %>%
  mutate(year = as.factor(year)) %>%
  group_by(year,zip,address) %>%
  summarise(amount=sum(amount))

full_by_year <- agrar_full %>%
  select(year,amount,land_based,is_firm,settlement_type)
full_by_year <- as.data.frame(full_by_year)

distinct_years <- full_by_year %>%
  select(year) %>%
  distinct()
years <- dim(distinct_years)[1]

```

```{r include=FALSE, echo = FALSE, message = FALSE, warning = FALSE}
# helper functions
save_3d_chart <- function(data, bins, target_var, dataset) {

  chart_data <- data %>%
    group_by(year) %>%
    mutate(amount = amount/1000000) %>%
    mutate(rank = ntile(amount,bins)) %>%
    group_by(year,rank) %>%
    summarise(cnt=n(),
              log_avg=log(mean(amount)),
              avg=mean(amount),
              log_sum=log(sum(amount)),
              sum=sum(amount)) %>%
    arrange(desc(year)) %>%
    reshape2::dcast(rank ~ year, value.var = target_var) %>%
    column_to_rownames(var = "rank") %>%
    as.matrix()
  
  hist3D(x = 1:bins, y = 1:years, z = chart_data,
         bty = "g", phi = 20, theta = -60,
         main = paste0(target_var, " (HUF mm)"),
         xlab = "bins", ylab = "years", zlab = target_var,
         border = "black", shade = 0.3, expand = 0.9, contour = TRUE,
         col=ramp.col(c("blue", "yellow", "red")),
         space = 0.25, d = 3,
         colkey = list(side = 4, length = 0.6, width = 0.5, dist = -0.04)
         )
  
  # save image
  png(filename=paste0(data_out, "3d_",dataset,"_",target_var,"_",bins,".png"))
  plotdev()
  dev.off()
}

pagebreak <- function() {
  if (knitr::is_latex_output())
    return("\\newpage")
  else
    return('<div style="page-break-before: always;" />')
}
```

# Problem Definition

Agricultural subsidies represent a huge section of the European Union's budget. The way this money is distributed is often inefficient and sometimes fraudulent as covered thoroughly by the [New York Times](https://www.nytimes.com/hu/2019/11/03/world/europe/eu-agrar-tamogatas-magyarorszag.html) and various Hungarian news sites ([here](https://index.hu/belfold/2016/04/15/igy_nez_ki_felcsut_kornyeke_bevasaroltak_orbanek_baratai/), [here](https://k.blog.hu/2016/04/15/agrar2015) or [here](https://index.hu/gazdasag/2016/07/28/en_onoknel_ki_a_helyi_oliharga/)). The amount of money distributed to Hungarian farmers is 1.5-2% of the country's GDP so it's no surprise that many people want to get a share of it. By analyzing the data, we could find interesting patterns about how this money is distributed.

Below are charts to show the proportion of subsidies funded by the EU and the Hungarian government.

```{r dodge-st, fig.show = "hold", out.width = "50%"}
ggplot(data = sum_by_dummies, aes(x = year, y = amount/1000000000, fill = source)) + 
  geom_bar(stat = "identity") +
  ylab("sum of total amount (bn HUF)") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  theme_minimal() +
  scale_fill_viridis(discrete=TRUE,direction=1,option = "D")
ggsave(paste0(data_out, "2d_sum_source.png"))

ggplot(data = sum_by_dummies, aes(x = year, y = amount/1000000000, fill = source)) + 
  geom_bar(stat = "identity", position = 'fill') +
  ylab("Proportion of total amount") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  theme_minimal() +
  scale_fill_viridis(discrete=TRUE,direction=1,option = "D")
ggsave(paste0(data_out, "2d_sum_source_stacked.png"))
```

The data is freely available on the website of the [Hungarian Treasury](https://www.mvh.allamkincstar.gov.hu/kozzeteteli-listak1) but it's difficult to work with. Most of the money comes from the EU so simply tampering with the data is not an option because it would be easy to spot this. Very limited analysis has been done on this data except for some high-level statistics and targeted investigations for high-profile actors but even that shows that the numbers are quite accurate. Similarly to other datasets made public by government agencies, this one is difficult to work with for various reasons. To hide the true nature of their activities, state actors often comply with their own laws only to a minimal extent. In most cases they will make public whatever they are legally required to do so, but they are not motivated to make it easy to work with. Another important reason for the badness of the data is that state agencies don’t represent an attractive workplace for highly skilled workers. As a result, government agencies simply lack the talent to make great, public datasets. Cleaning and enriching the dataset would allow investigative journalists, non-profits or curious citizens to learn about this domain. This paper covers the initial efforts to solve this problem. The data transformation and analysis scripts are all publicly available on [GitHub](https://github.com/gergo/hungarianAgricultureSubsidies).

`r pagebreak()`
# Data Processing Pipeline

The data has to go through a cleaning and transformation pipeline for various error corrections and enhancements to be applied. There are other, freely available datasets (e.g.: socio-economic, elections) that can be merged with the agricultural subsidies data to provide a wide range of descriptive variables. To eliminate redundancies and make the dataset more compact we can normalize it by extracting various aspects of the data.

When creating the data engineering pipeline the aim was to make it as automatic as possible. Anyone who has the project checked out from the [GitHub repository](https://github.com/gergo/hungarianAgricultureSubsidies) and has the runtime dependencies in place should be able to build the project from scratch. Data processing and analysis can be performed on a personal computer using open-source or free tools.

## Runtime dependencies
- **bash**: The various data sources can be downloaded and processed by executing a handful of shell scripts. For this a unix environment wiht bash is required.
- **KDB**: To efficiently process the raw data and merge it with other datasets I use a programming language called q that comes with an in-memory database solution called KDB. It's free to [download](https://kx.com/connect-with-us/download/).
- **Python 3**: The project contains a script to call the Google Maps API to geocode addresses. This creates a standardized format for the addresses and also adds latitude and longitude coordinates to the addresses of the beneficiaries.
- **R**: The data analysis is done in R so an [R runtime](https://www.r-project.org/) is also needed.

## Downlad raw data

The raw dataset is available from 2011 on the website of the [Hungarian Treasury](https://www.mvh.allamkincstar.gov.hu/kozzeteteli-listak1). The site provides an on-line search interface and links to compressed csv files that contain yearly batches of subsidies. The website has direct links to the datasets from the past 5 years but based on the patterns it is not difficult to figure out how to download earlier batches. Data for 2010 used to be available on the site but the link no longer works so I added that as part of the project and adjusted the download scripts to handle that differently. The raw files are quite large:

```{bash}
ls -lh ../input/csv/*.csv | awk '{print $5,$9}' | paste - -
```

## Fix character encoding

The original files are in ISO-8859-2 encoding (aka. Latin-2). After converting the data to UTF8 the contents look like this:

```{bash}
# most rows are long; to make the sample fit to the generated document I filter for rows with <100 characters
cat ../input/csv/utf8_2011.csv | head -n1000 | awk 'length($0) < 100' | head -n6
```

The data is now in a format that programming languages can deal with.

## Basic data cleaning
During exploratory data analysis I found multiple versions of the same entity. Fixing this is important if we want to get an accurate view on at aggregate subsidies by beneficiaries.
 - *White spaces*:  Some names contained multiple spaces instead of a single space when separating name parts. I replaced all whitespace group with a single space in the names.
 - *Character cases*: Some names had multiple versions with inconsistent capitalizations (eg. all caps) so I split names by spaces and capitalized only the first character in each name part.
 - *Remove dots*: Some names had multiple variations based on how abbreviations in their name parts were written: some had '.', other rows did not.

The above steps reduced the unique number of entities from 249087 to 248245.

## Find firms

Firms and individuals are not separated in the data. I made the initial assumption that any name that has 8 or more name parts is a firm. For names that have 8 or fewer name parts I identified a large number of keywords that indicate that the beneficiari is not an individual. These are not just firms but churches, municipalities, clubs and all sorts of other organizations but for simplicity I call these firms.

```{bash}
paste -s -d";" ../input/names/firm_keywords.txt
```

The proportion of subsidies received by firms and individuals can be seen in the below charts.

```{r fig.show = "hold", out.width = "50%"}
ggplot(data = sum_by_dummies, aes(x = year, y = amount/1000000000, fill = is_firm)) + 
  geom_bar(stat = "identity") +
  labs(y = "sum amount (bn HUF)") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  theme_minimal() +
  scale_fill_viridis(discrete=TRUE,direction=1,option = "D")
ggsave(paste0(data_out, "2d_sum_isfirm.png"))

ggplot(data = sum_by_dummies, aes(x = year, y = amount/1000000000, fill = is_firm)) + 
  geom_bar(stat = "identity", position = 'fill') +
  labs(y = "Proportion by group") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  theme_minimal() +
  scale_fill_viridis(discrete=TRUE,direction=1,option = "D")
ggsave(paste0(data_out, "2d_sum_isfirm_stacked.png"))
```

## Gender Information
Most of the beneficiaries are Hungarian citizens. The legal given names are available on the [internet](http://www.nytud.mta.hu/oszt/nyelvmuvelo/utonevek/). Some female citizens changed their legal name to that of their husbands and a "-né" postfix. This pattern is easy to find so I took this into account. The distribution of male and female beneficiaries is in the below charts.

```{r fig.show = "hold", out.width = "50%"}
by_gender <- agrar_full %>%
  filter(is_firm == 0) %>%
  select(year,amount,land_based,settlement_type,gender) %>%
  mutate(year = as.factor(year),
         gender = as.factor(gender)) %>%
  mutate(gender = forcats::fct_explicit_na(gender, na_level = "unknown")) %>%
  group_by(year,land_based,settlement_type,gender) %>%
  summarise(amount=sum(amount))

ggplot(data = by_gender, aes(x = year, y = amount/1000000000, fill = gender)) + 
  geom_bar(stat = "identity") +
  labs(y = "sum amount (bn HUF)") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  theme_minimal() +
  scale_fill_viridis(discrete=TRUE,direction=1,option = "D")
ggsave(paste0(data_out, "2d_sum_by_gender.png"))

ggplot(data = by_gender, aes(x = year, y = amount/1000000000, fill = gender)) + 
  geom_bar(stat = "identity", position = "fill") +
  ylab("Proportion of total amount") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  theme_minimal() +
  scale_fill_viridis(discrete=TRUE,direction=1,option = "D")
ggsave(paste0(data_out, "2d_sum_by_gender_stacked.png"))
rm(by_gender)
```

## Clean Addresses

Zip codes, settlement parts, county capitals, ksh_id and basic stats from ksh (http://www.ksh.hu/docs/helysegnevtar/hnt_letoltes_2019.xls)
Some missing zip codes using Hungarian Postal Service (https://www.posta.hu/static/internet/download/Iranyitoszam-Internet_uj.xlsx)
Clean Addresses
330 000 unique addresses with typos, formatting differences.
use [Google Maps Geocoding Api](https://maps.googleapis.com/maps/api/geocode/) to standardize addresses and get latitude/longitude coordinates  
Manual mapping files

`r pagebreak()`
# Data Exploration

## Descriptive Stats

```{r include=FALSE}
stat_controls <- tableby.control(
  test = FALSE,
  total = FALSE,
  numeric.stats = c("meansd", "q1q3", "medianq1q3", "range"),
  cat.stats = c("countpct"),
  stats.labels = list(
    N='Count',
    meansd = "Mean (SD)",
    medianq1q3 = "Median (Q1, Q3)",
    q1q3 = "Q1, Q3",
    range = "Min - Max"
  )
)

full_stat_table <- tableby(year ~ .,
                     data = agrar_full %>%
                       select(year,amount) %>%
                       mutate(amount = amount / 1000),
                     control = stat_controls)

by_address_data <- sum_by_address %>%
                    ungroup() %>%
                    select(year,amount) %>%
                    mutate(amount = amount / 1000)

by_address_stat_table <- tableby(year ~ .,
                       data = by_address_data,
                        control = stat_controls)
rm(by_address_data)
```

Let's see some descriptive statistics of the whole dataset.

```{r}
##########################
amount <- agrar_full$amount
# Save the mean, sd, min, max, quantiles of won amounts
descr_amount <- as.matrix( c( mean(amount) , sd(amount), min(amount) , max(amount) , quantile(amount,.50), quantile(amount,.95) , length(amount) ) )
# Name the dimensions as a list to have neat output 
dimnames(descr_amount)[[1]] <- list('mean','sd','min','max', 'p50' , 'p95' , 'n')
print(descr_amount)
rm(amount)


```

```{r results="asis"}
summary(full_stat_table,
        title = "Yearly statistics of all agricultural subsidies (1000s HUF)",
        digits=0, digits.p=0, digits.pct=0,digits.n=NA)
```

```{r results="asis"}
summary(by_address_stat_table,
        title = "Agricultural subsidies summed by beneficiaries (1000s HUF)",
        digits=0, digits.p=0, digits.pct=0,digits.n=NA)
```


## Settlement type

Let's see how subsidies are distributed among settlement types. It is important to note that the address is the official address of the firm or individual who received the money, not the address of the land, for which the money was given. Most of the actual land belongs to villages and small towns but the proporion of the subsidies received by entities who are based in the country capital or county capitals is quite significant.

```{r fig.show = "hold", out.width = "50%"}
ggplot(data = sum_by_dummies, aes(x = year, y = amount/1000000000, fill = settlement_type)) + 
  geom_bar(stat = "identity") +
  labs(y = "sum amount (bn HUF)") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  theme_minimal() +
  scale_fill_viridis(discrete=TRUE,direction=1,option = "D")
ggsave(paste0(data_out, "2d_sum_settlement_type.png"))

ggplot(data = sum_by_dummies, aes(x = year, y = amount/1000000000, fill = settlement_type)) + 
  geom_bar(stat = "identity", position = 'fill') +
  labs(y = "Proportion of subsidies") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  theme_minimal() +
  scale_fill_viridis(discrete=TRUE,direction=1,option = "D")
ggsave(paste0(data_out, "2d_sum_settlement_type_stacked.png"))
```

Let's see the average amounts by decil for each year.

```{r fig.show = "hold", out.width = "50%", echo=FALSE, results='hide', fig.height=6}
save_3d_chart(full_by_year, 10, "log_avg", "all_yearly")
save_3d_chart(full_by_year, 10, "avg", "all_yearly")
```

```{r fig.show = "hold", out.width = "50%", echo=FALSE, results='hide', fig.height=6}
save_3d_chart(full_by_year, 10, "log_sum", "all_yearly")
save_3d_chart(full_by_year, 10, "sum", "all_yearly")
```


## Overall distribution

```{r include=FALSE}
sum_by_address <- agrar_full %>%
  select(year,zip,address,amount) %>%
  mutate(year = as.factor(year)) %>%
  group_by(year,zip,address) %>%
  summarise(amount=sum(amount))

sum_by_address_decils <- sum_by_address %>%
  group_by(year) %>%
  mutate(rank = ntile(amount,10)) %>%
  mutate(rank = as.factor(rank)) %>%
  mutate(rank = fct_reorder(rank, desc(rank))) %>%
  group_by(year,rank) %>%
  summarise(sum=sum(amount)) %>%
  arrange(desc(year))
```

```{r fig.show = "hold", out.width = "50%", echo=FALSE, results='hide', fig.height=6}
save_3d_chart(sum_by_address, 10, "avg", "summed-by-address")
save_3d_chart(sum_by_address, 10, "log_avg", "summed-by-address")
```

```{r fig.show = "hold", out.width = "50%", echo=FALSE, results='hide', fig.height=6}
save_3d_chart(sum_by_address, 30, "avg", "summed-by-address")
save_3d_chart(sum_by_address, 30, "log_avg", "summed-by-address")
```

```{r fig.show = "hold", out.width = "50%", echo=FALSE, results='hide', fig.height=6}
save_3d_chart(sum_by_address, 10, "sum", "summed-by-address")
save_3d_chart(sum_by_address, 10, "log_sum", "summed-by-address")
```

```{r fig.show = "hold", out.width = "50%", echo=FALSE, results='hide', fig.height=6}
save_3d_chart(sum_by_address, 30, "sum", "summed-by-address")
save_3d_chart(sum_by_address, 30, "log_sum", "summed-by-address")
```


```{r fig.show = "hold", out.width = "50%"}
ggplot(data = sum_by_address_decils, aes(x = year, y = sum/1000000000, fill = rank)) + 
  geom_bar(stat = "identity") +
  labs(y = "sum amount (HUF bn)") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  theme(axis.text.y = element_text(angle = 90, hjust = 1)) +
  theme_minimal() +
  scale_fill_viridis(discrete=TRUE,direction=1,option = "D")
ggsave(paste0(data_out, "2d_sum_by_decils.png"))

ggplot(data = sum_by_address_decils, aes(x = year, y = sum/1000000000, fill = rank)) + 
  geom_bar(stat = "identity", position = 'fill') +
  labs(y = "Pct of total amount") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  theme_minimal() +
  scale_fill_viridis(discrete=TRUE,direction=1,option = "D")
ggsave(paste0(data_out, "2d_sum_by_decils_stacked.png"))
```


```{r include=FALSE}
rm(sum_by_address_decils)
rm(sum_by_address)
```

## Overall distribution

```{r include=FALSE}
sum_by_address_indiv <- agrar_full %>%
  filter(is_firm == 0) %>% 
  select(year,zip,address,amount) %>%
  mutate(year = as.factor(year)) %>%
  group_by(year,zip,address) %>%
  summarise(amount=sum(amount))

sum_by_address_indiv_decils <- sum_by_address_indiv %>%
  group_by(year) %>%
  mutate(rank = ntile(amount,10)) %>%
  mutate(rank = as.factor(rank)) %>%
  mutate(rank = fct_reorder(rank, desc(rank))) %>%
  group_by(year,rank) %>%
  summarise(sum=sum(amount)) %>%
  arrange(desc(year))
```


```{r fig.show = "hold", out.width = "50%", echo=FALSE, results='hide', fig.height=6}
save_3d_chart(sum_by_address_indiv, 10, "avg", "summed-by-address-indiv")
save_3d_chart(sum_by_address_indiv, 10, "log_avg", "summed-by-address-indiv")
```
```{r fig.show = "hold", out.width = "50%", echo=FALSE, results='hide', fig.height=6}
save_3d_chart(sum_by_address_indiv, 10, "sum", "summed-by-address-indiv")
save_3d_chart(sum_by_address_indiv, 10, "log_sum", "summed-by-address-indiv")
```

```{r fig.show = "hold", out.width = "50%"}
ggplot(data = sum_by_address_indiv_decils, aes(x = year, y = sum/1000000000, fill = rank)) + 
  geom_bar(stat = "identity") +
  labs(y = "sum amount (HUF bn)") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  theme(axis.text.y = element_text(angle = 90, hjust = 1)) +
  theme_minimal() +
  scale_fill_viridis(discrete=TRUE,direction=1,option = "D")
ggsave(paste0(data_out, "2d_sum_by_indiv.png"))

ggplot(data = sum_by_address_indiv_decils, aes(x = year, y = sum/1000000000, fill = rank)) + 
  geom_bar(stat = "identity", position = 'fill') +
  labs(y = "Pct of total amount") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  theme_minimal() +
  scale_fill_viridis(discrete=TRUE,direction=1,option = "D")
ggsave(paste0(data_out, "2d_sum_by_indiv_stacked.png"))
```

```{r include=FALSE}
rm(sum_by_address_indiv)
rm(sum_by_address_indiv_decils)
```




## Land-based distribution

```{r include=FALSE}
############################
# land-based trends
############################
full_by_year_by_type <- agrar_full %>%
  filter(land_based == 1) %>%
  select(amount,year,reason) %>%
  group_by(year,reason) %>%
  summarise(amount=sum(amount)) %>%
  as.data.frame()
```

```{r fig.show = "hold", out.width = "50%"}
ggplot(data = full_by_year_by_type, aes(x = year, y = amount/1000000000, fill = reason)) + 
  geom_bar(stat = "identity") +
  ylab("percent of total amount") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  theme_minimal() +
  scale_fill_viridis(discrete=TRUE,direction=1,option = "D")
ggsave(paste0(data_out, "2d_land_based_trends.png"))

ggplot(data = full_by_year_by_type, aes(x = year, y = amount/1000000000, fill = reason)) + 
  geom_bar(stat = "identity", position = "fill") +
  ylab("percent of total amount") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  theme_minimal() +
  scale_fill_viridis(discrete=TRUE,direction=1,option = "D")
ggsave(paste0(data_out, "2d_land_based_trends_stacked.png"))
```



```{r fig.show = "hold", out.width = "50%"}
ggplot(data = sum_by_dummies, aes(x = year, y = amount/1000000000, fill = land_based)) + 
  geom_bar(stat = "identity") +
  labs(y = "sum amount (bn HUF)") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  theme_minimal() +
  scale_fill_viridis(discrete=TRUE,direction=1,option = "D")
ggsave(paste0(data_out, "2d_sum_land_based.png"))

ggplot(data = sum_by_dummies, aes(x = year, y = amount/1000000000, fill = land_based)) + 
  geom_bar(stat = "identity", position = 'fill') +
  labs(y = "sum amount (bn HUF)") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  theme_minimal() +
  scale_fill_viridis(discrete=TRUE,direction=1,option = "D")
ggsave(paste0(data_out, "2d_sum_land_based_stacked.png"))
```

```{r include=FALSE}
rm(full_by_year_by_type)  
rm(sum_by_dummies)
```

## Individual trends

```{r include=FALSE}
###################################
# sum_by_address - individuals
###################################

sum_by_address_indiv_land <- agrar_full %>%
  filter(is_firm == 0, land_based == 1) %>% 
  select(year,zip,address,amount) %>%
  mutate(year = as.factor(year)) %>%
  group_by(year,zip,address) %>%
  summarise(amount=sum(amount))

sum_by_address_indiv_land_decils <- sum_by_address_indiv_land %>%
  group_by(year) %>%
  mutate(rank = ntile(amount,10)) %>%
  mutate(rank = as.factor(rank)) %>%
  mutate(rank = fct_reorder(rank, desc(rank))) %>%
  group_by(year,rank) %>%
  summarise(sum=sum(amount)) %>%
  arrange(desc(year))
```

```{r fig.show = "hold", out.width = "50%", echo=FALSE, results='hide', fig.height=6}
save_3d_chart(sum_by_address_indiv_land, 10, "avg", "summed-by-address-indiv")
save_3d_chart(sum_by_address_indiv_land, 10, "log_avg", "summed-by-address-indiv")
```

```{r fig.show = "hold", out.width = "50%", echo=FALSE, results='hide', fig.height=6}
save_3d_chart(sum_by_address_indiv_land, 10, "sum", "summed-by-address-indiv")
save_3d_chart(sum_by_address_indiv_land, 10, "log_sum", "summed-by-address-indiv")
```

```{r fig.show = "hold", out.width = "50%"}
ggplot(data = sum_by_address_indiv_land_decils, aes(x = year, y = sum/1000000000, fill = rank)) + 
  geom_bar(stat = "identity") +
  labs(y = "sum amount (HUF bn)") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  theme(axis.text.y = element_text(angle = 90, hjust = 1)) +
  theme_minimal() +
  scale_fill_viridis(discrete=TRUE,direction=1,option = "D")
ggsave(paste0(data_out, "2d_sum_by_indiv_land.png"))

ggplot(data = sum_by_address_indiv_land_decils, aes(x = year, y = sum/1000000000, fill = rank)) + 
  geom_bar(stat = "identity", position = 'fill') +
  labs(y = "Pct of total amount") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  theme_minimal() +
  scale_fill_viridis(discrete=TRUE,direction=1,option = "D")
ggsave(paste0(data_out, "2d_sum_by_indiv_land_stacked.png"))
```

```{r include=FALSE}
rm(sum_by_address_indiv_land)
rm(sum_by_address_indiv_land_decils)
```

# Modeling

Is there a relationship between the amount of money given to a region and the political affiliation of the municipality/county?

Election data: https://drive.google.com/file/d/1E8oMR97J-_H9vnymzEdF29GN0o6LlkqN/

Additional variables from KSH (http://statinfo.ksh.hu/Statinfo/index.jsp) 


